---
title: "Structural Breaks in HS6 hs6_codes"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: fill
runtime: shiny
---

```{r global, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```


```{r}
## Loading the packages
library(pacman)
p_load(tidyverse, dplyr, ggplot2, scales, plotly, flexdashboard, tidyr, shiny, rsconnect, readxl, readr, strucchange, kableExtra)

```

```{r}
## Loading the necessary data
#setwd("C:/Users/Ciani/Documents/Disruptive Technologies/Shiny App")

igpc_data <- read_csv("data/dynamic_igpc_results_normalised.csv", show_col_types = FALSE)
trade_data <- read_csv("data/clean_trade_data.csv", show_col_types = FALSE) 
patent_data <- read.csv("data/patent_data.csv", sep=";", quote="")
ipc_hs_conversion_table <- read_csv("data/ipc_hs_conversion_table.csv", show_col_types = FALSE)
ipc_definitions <- read_excel("data/IPC_Definitions.xlsx")


```


```{r}

### I just remove all of these options. 

igpc_data <- igpc_data %>%
  select(-c(hs_code, gts, tc, k)) %>%
  mutate(
    hs6_code = as.numeric((hs6_code)))

trade_data <- trade_data %>%
  select(-c(k)) %>%
  mutate(
    hs6_code = as.numeric(hs6_code)
  )

## Remove all those quotes
patent_data[] <- lapply(patent_data, function(x) {
  if (is.character(x)) gsub('"', "", x) else x
})

```

```{r}
## Getting choice options for igpc

igpc_data <- igpc_data %>%
  mutate(
    year = as.integer(year),
    description = as.character(description)
  ) %>%
  arrange(description, year)

choice_igpc_df <- igpc_data %>% distinct(hs6_code, description)
choice_igpc <- setNames(choice_igpc_df$hs6_code,
                        paste0(choice_igpc_df$description, " — ", choice_igpc_df$hs6_code))


```

```{r}

## Cleaning the trade data

choices_baci_df <- trade_data %>%
  distinct(hs6_code, description) %>%
  arrange(description)

labels_val <- paste0(choices_baci_df$description, " — ", choices_baci_df$hs6_code)
choice_baci <- setNames(choices_baci_df$hs6_code, labels_val)

```


```{r}

## Cleaning the definitions

ipc_definitions <- ipc_definitions %>%
  mutate(ipc_definition = tolower(paste(IPC3_definition, IPC4_definition, sep = " - "))) %>%
  select(IPC4, ipc_definition)

# Join HS6–IPC map with definitions
hs_ipc_final <- ipc_hs_conversion_table %>%
  mutate(code = as.character(code), ipc4 = toupper(ipc4)) %>%
  left_join(ipc_definitions, by = c("ipc4" = "IPC4"))

# Make sure patent_data columns line up (rename if your columns differ)
names(patent_data)[1] <- "year"

patent_data <- patent_data %>%
  rename(ipc4 = cpc4, patents = n_applications) %>%   # adjust if needed
  mutate(
    ipc4 = toupper(trimws(ipc4)),
    year = as.integer(year),
    patents = as.numeric(patents)
  )

## choices

choice_patents <- igpc_data %>%
  distinct(hs6_code, description) %>%
  arrange(description) %>%
  mutate(label = paste0(description, " — ", hs6_code)) %>%
  with(setNames(hs6_code, label))

```

# Product Centrality

## Sidebar {.sidebar}

```{r}

selectizeInput(
  inputId = "hs6_code",
  label   = "Search or select a Good:",
  choices = NULL,                 # <- populated from server
  options = list(
    placeholder     = 'Type keywords (e.g., "lithium battery")',
    closeAfterSelect = TRUE
  )
)

## Populate choices from the server
observe({
  updateSelectizeInput(
    session,
    "hs6_code",
    choices  = choice_igpc,    
    selected = choice_igpc[[1]],
    server   = TRUE               
  )
})
```

For this Dashboard, we constructed the Integrated Global Product Centrality Indicator from Fetzer, Lambert, Feld, Garg (2024) based on CEPII's BACI trade data. 

The IGPC has been normalised at 100 in 2002 for all goods. 

The data spans from 2002 to 2023. 

One can search for any HS6 good either by hs6_code or description.

Source: Fetzer, year., Lambert, P. J., Feld, B., & Garg, P. (2024). AI-generated production networks: Measurement and applications to global trade.

## Column 1 

### Centrality Over Time {data-height=650} 

```{r, fill = TRUE}

plotlyOutput("lineplot")  

```

```{r}

output$lineplot <- renderPlotly({
  req(input$hs6_code)

  filtered_data <- igpc_data %>%
    mutate(hs6_code = as.character(hs6_code)) %>%
    filter(hs6_code == as.character(input$hs6_code)) %>%
    arrange(year) %>%
    transmute(
      year       = as.numeric(year),
      IGPC_index = as.numeric(IGPC_index)
    ) %>%
    drop_na()

  validate(
    need(nrow(filtered_data) > 0, "No data available for the selected HS6 Code")
  )

  plot_ly(
    data = filtered_data,
    x    = ~year,
    y    = ~IGPC_index,
    type = "scatter",
    mode = "lines+markers",
    line = list(
      color = "#4169C3",     # line color
      width = 3              # line thickness
    ),
    marker = list(
      color = "#194ABB",     # marker (dot) color
      size  = 8             # marker size
    ),
    hovertemplate = "Year: %{x}<br>IGPC: %{y}<extra></extra>"
  ) %>%
    layout(
      title = "Centrality in Global Trade",
      xaxis = list(title = "Year"),
      yaxis = list(title = "Integrated Global Product Centrality")
    )
})

```

### Structural Break test {data-height=350}

```{r}

htmlOutput("break_table")

```

```{r}
output$break_table <- renderUI({
  req(input$hs6_code)

  df <- igpc_data %>%
    filter(hs6_code == input$hs6_code) %>%
    arrange(year) %>%
    select(year, IGPC_index) %>%
    drop_na()

  # Need at least 10-15 observations for meaningful break test
  validate(
    need(nrow(df) > 15, "Not enough data for break test (need at least 15 observations)")
  )
  
  # Set minimum segment size (h parameter)
  n <- nrow(df)
  h <- max(3, ceiling(0.15 * n))  # At least 3 observations per segment, or 15% of data
  
  # Make sure h is valid
  validate(
    need(h < (n - h), "Series too short for structural break analysis")
  )

  # SupF test with from parameter
  test <- sctest(IGPC_index ~ year, type = "supF", data = df, from = h/n)
  stat <- unname(test$statistic)
  pval <- unname(test$p.value)
  sig  <- pval < 0.05

  # Estimated break year(s) with explicit h
  bp    <- breakpoints(IGPC_index ~ year, data = df, h = h)
  breaks <- df$year[bp$breakpoints]
  breaks <- breaks[!is.na(breaks)]
  breaks_str <- if (length(breaks)) paste(breaks, collapse = ", ") else "—"

  tab <- tibble::tibble(
    `Structural Break Test` = c("Test", "F statistic", "p-value", "Significant (5%)", "Estimated break year(s)"),
    Value = c("supremum F test", sprintf("%.3f", stat), scales::pvalue(pval, accuracy = 0.001),
              ifelse(sig, "Yes", "No"), breaks_str)
  )

  kb <- knitr::kable(tab, format = "html", align = c("l", "r"), escape = TRUE) |>
    kableExtra::kable_styling(full_width = TRUE,
                              bootstrap_options = c("condensed", "striped", "hover"),
                              font_size = 15 )

  if (sig) {
    kb <- kb |> kableExtra::row_spec(4, bold = TRUE)
  }

  htmltools::HTML(kb)
})

```

# Trade Data

## Sidebar {.sidebar}

```{r}
selectizeInput(
  inputId = "hs6_code_val",
  label   = "Search HS6 good:",
  choices = NULL,
  options = list(
    placeholder = 'Type keywords or a hs6_code…',
    closeAfterSelect = TRUE
  )
)

observe({
  updateSelectizeInput(
    session,
    "hs6_code_val",
    choices  = choice_baci,
    selected = choice_baci[[1]],
    server   = TRUE
  )
})

```


In this case, we have analysed trade value of HS6 goods compared to the average trade value in the HS4 category. 

For this Dashboard, we used CEPII's BACI trade data. 

The trade value has been normalised at 100 in 2002 for all goods. 

The data spans from 2002 to 2023. 

One can search for any HS6 good either by hs6_code or description.

## Column 1

### Value of Trade {data-height=650}

```{r}

plotlyOutput("lineplot_val", height = "100%")

```

```{r}
output$lineplot_val <- renderPlotly({
  req(input$hs6_code_val)

  # Ensure consistent types
  df <- trade_data %>%
    mutate(hs6_code = as.character(hs6_code)) %>%
    filter(hs6_code == as.character(input$hs6_code_val)) %>%
    arrange(year) %>%
    drop_na(value_normalised, hs4_value_average)

  validate(
    need(nrow(df) > 0, "No trade data available for this HS6 code.")
  )

  # Create long format using base R (no pivot_longer coercion issues)
  plt_df <- data.frame(
    year = rep(df$year, 2),
    series = factor(
      rep(c("HS6 (2002=100)", "HS4 avg (2002=100)"), each = nrow(df)),
      levels = c("HS6 (2002=100)", "HS4 avg (2002=100)")
    ),
    index = c(df$value_normalised, df$hs4_value_average)
  )

  # Subtitle
  matched_names <- names(choice_baci)[choice_baci == input$hs6_code_val]
  sub <- if (length(matched_names) > 0) matched_names[1] else paste("HS6:", input$hs6_code_val)

  title_html <- paste0(
    "HS6 vs HS4 Average — Trade Value (2004 = 100)",
    "<br><span style='font-size:12px;'>", htmltools::htmlEscape(sub), "</span>"
  )

  # Pure plotly (no ggplotly)
plot_ly() %>%
  add_trace(
    data = plt_df[plt_df$series == "HS6 (2002=100)",],
    x = ~year,
    y = ~index,
    type = "scatter",
    mode = "lines+markers",
    name = "HS6 (2002=100)",
    line = list(color = "#4169C3", width = 4),
    marker = list(color = "#194ABB", size = 8)
  ) %>%
  add_trace(
    data = plt_df[plt_df$series == "HS4 avg (2002=100)",],
    x = ~year,
    y = ~index,
    type = "scatter",
    mode = "lines+markers",
    name = "HS4 avg (2002=100)",
    line = list(color = "#C8CFDE", width = 4),
    marker = list(color = "#9BADD5", size = 8)
  ) %>%
    layout(
      title = list(text = title_html),
      xaxis = list(title = "Year"),
      yaxis = list(title = "Trade value index (2004 = 100)"),
      legend = list(orientation = "h"),
      margin = list(l = 60, r = 20, b = 40, t = 80)
    )
})

```

### Structural Breaks {data-height=350}

```{r}

htmlOutput("break_table_val")

```

```{r}
output$break_table_val <- renderUI({
  req(input$hs6_code)

  df <- trade_data %>%
    filter(hs6_code == input$hs6_code) %>%
    arrange(year) %>%
    select(year = year, value_normalised) %>%
    drop_na()

  # Need enough data
  validate(need(nrow(df) > 8, "Not enough data for break test (need > 8 observations)."))

  n <- nrow(df)
  k <- 2  # intercept + year
  h <- max(k + 1, ceiling(0.20 * n))  # e.g. 20% window, and strictly > k

  # If even that won'year work, stop with a clear message
  validate(need(h < n, sprintf(
    "Series too short for structural break test (n=%d, required minimum segment size=%d).", n, h
  )))

  # SupF test with matching 'from' = h/n
  test <- sctest(value_normalised ~ year, type = "supF", data = df, from = h / n)
  stat <- unname(test$statistic)
  pval <- unname(test$p.value)
  sig  <- pval < 0.05

  # Breakpoints with explicit h
  bp      <- breakpoints(value_normalised ~ year, data = df, h = h)
  breaks  <- df$year[bp$breakpoints]
  breaks  <- breaks[!is.na(breaks)]
  breaks_str <- if (length(breaks)) paste(breaks, collapse = ", ") else "—"

  tab <- tibble::tibble(
    `Structural Break Test (Trade, HS6 index)` =
      c("Test", "F statistic", "p-value", "Significant (5%)", "Estimated break year(s)", "Min segment size (h)"),
    Value = c("supremum F test",
              sprintf("%.3f", stat),
              scales::pvalue(pval, accuracy = 0.001),
              ifelse(sig, "Yes", "No"),
              breaks_str,
              h)
  )

  kb <- knitr::kable(tab, format = "html", align = c("l","r"), escape = TRUE) |>
    kableExtra::kable_styling(full_width = TRUE,
                              bootstrap_options = c("condensed","striped","hover"),
                              font_size = 15)
  if (sig) kb <- kb |> kableExtra::row_spec(4, bold = TRUE)

  htmltools::HTML(kb)
})

```


# Patent Data

## Sidebar {.sidebar}


```{r}
# Sidebar for patent tab
selectizeInput(
  inputId = "hs6_code_pat",
  label   = "Search HS6 good:",
  choices = NULL,
  options = list(
    placeholder = "Type HS6 code…",
    closeAfterSelect = TRUE
  )
)

## Server
observe({
  updateSelectizeInput(
    session,
    "hs6_code_pat",
    choices  = choice_patents,
    selected = choice_patents[[1]],
    server   = TRUE
  )
})

```

Select the HS6 Code to see the number of patents that are published in the most relevant field and the list of all the relevant fields. 

Source for the HS-IPC links: "Getting patents and economic data to speak to each other: An ‘algorithmic links with probabilities’ approach for joint analyses of patenting and economic activity." Nikolas ZOlas and Travis Lybbert, Research Policy, 43.3 (2014): 530-542.

Source for Patent data: PATSTAT

## Column 1 

### Inventions and Patents Over Time {data-height=650} 


```{r}

plotlyOutput("patent_line", height = "100%")

```


```{r}
## Here we show the plot
output$patent_line <- renderPlotly({
  req(input$hs6_code_pat)

  prob_col <- "probability_weight"

  # Normalize hs6_code type
  hs_ipcs <- hs_ipc_final %>%
    mutate(hs6_code = as.character(code)) %>%
    filter(hs6_code == as.character(input$hs6_code_pat)) %>%
    arrange(desc(.data[[prob_col]]))

  validate(need(nrow(hs_ipcs) > 0, "No IPC mapping for this HS6 code."))

  # Select top IPC4 class
  top_ipc <- hs_ipcs %>% slice(1) %>% pull(ipc4)

  # Patent time series for selected IPC4
  ser <- patent_data %>%
    mutate(ipc4 = as.character(ipc4)) %>%     # ensure matching
    filter(ipc4 == top_ipc) %>%
    arrange(year) %>%
    transmute(
      year    = as.numeric(year),
      patents = as.numeric(patents)
    ) %>%
    drop_na()

  validate(need(nrow(ser) > 0, "No patent data for this IPC class."))

  # Subtitle from HS6 code selection
  sub <- names(which(choice_patents == input$hs6_code_pat))[1]
  ttl <- paste0(
    "Patents (Most likely IPC4: ", top_ipc, ")",
    "<br><span style='font-size:12px;'>",
    htmltools::htmlEscape(sub),
    "</span>"
  )

  # Pure plotly output
  plot_ly(
    data = ser,
    x = ~year,
    y = ~patents,
    type = "scatter",
    mode = "lines+markers",
    line = list(color = "#3769aa", width = 3),
    marker = list(color = "#01438f", size = 8)
  ) %>%
    layout(
      title = list(text = ttl),
      xaxis = list(title = "Year"),
      yaxis = list(title = "Patent count"),
      margin = list(l = 60, r = 20, b = 40, t = 80)
    )
})

```

### All Relevant IPC Categories

```{r}
## Here we show the Kable Table with the most probable IPC Categories
htmlOutput("ipc_table")
```

```{r}
output$ipc_table <- renderUI({
  req(input$hs6_code_pat)

  prob_col <- "probability_weight"   # same caveat as above
  tbl <- hs_ipc_final %>%
    filter(code == input$hs6_code_pat) %>%
    select(IPC4 = ipc4, Probability = all_of(prob_col), Definition = ipc_definition) %>%
    arrange(desc(Probability))

  validate(need(nrow(tbl) > 0, "No IPC classes found for this HS6 hs6_code."))

  kb <- knitr::kable(tbl, format = "html", align = c("c","r","l"), escape = TRUE) %>%
    kableExtra::kable_styling(full_width = TRUE,
                              bootstrap_options = c("condensed","striped","hover"),
                              font_size = 14)

  htmltools::HTML(kb)
})

```

